---
title: "Property Analysis with Chi-Square Test"
author: "Caio di Felice Cunha"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business problem

The dataset describes properties (price, size, zip code, type, status, rent
status, city). The question is whether two of the categorical variables,
`Type_Property` and `Status_Property` (New vs Old), are related. The
appropriate test for a relationship between two categorical variables is the
chi-square test of independence.

H0: the two variables are independent (no relationship).
H1: the two variables are related.

We reject H0 when the p-value is below 0.05. The test assumes independent
observations and an expected count of at least 5 in every cell.

This report was knit from `Property-Analysis.Rmd`. It uses base R only and
requires R >= 4.0.

## Understand the data

```{r load}
df <- read.csv("data.csv")
str(df)
head(df)
dim(df)
```

```{r split}
# Variables under test
x <- df$Type_Property
y <- df$Status_Property

unique(x)
unique(y)
```

## Cross tables

```{r tables}
# Counts
table(x, y)

# Percentages
round(prop.table(table(x, y)) * 100, 2)
```

## Chi-square test (full dataset)

```{r test-full}
chisq.test(x, y)
chisq.test(table(x, y))
```

The test returns `X-squared = 868.75, df = 4, p-value < 2.2e-16`. The p-value is
far below 0.05, so we reject H0: property type and status are related.

## Exercise: exclude apartments

If apartments are removed, does the relationship hold?

```{r filter}
x <- droplevels(df$Type_Property[df$Type_Property != "Apartment"])
y <- droplevels(df$Status_Property[df$Type_Property != "Apartment"])

unique(x)
unique(y)

table(x, y)
round(prop.table(table(x, y)) * 100, 2)
```

```{r test-noapt}
chisq.test(x, y)
chisq.test(table(x, y))
```

Now the test returns `X-squared = 0.79718, df = 3, p-value = 0.8501`. The
p-value is well above 0.05, so we do not reject H0: among non-apartment types
there is no detectable association.

## Conclusion

The overall association between property type and status is driven entirely by
apartments. Apartments are disproportionately New (990 of 3891) compared with
every other type (for example 7 of 364 houses with a backyard). Once apartments
are excluded the relationship disappears.
